---
title: "GxE"
editor: visual
---
 Introduction: Nowadays, climate change is a pretty serious phenomenon and that's why everyone is paying a lot of attention to it. climate change has significant impacts on various sectors, including animal production. climate change is affecting milk production, both the quantity of production and its composition. Heat stress negatively affects the cows' comfort and overall health. Spain is one of the countries that is most affected by this climate change, since its affect varies between the north and the south of this country. As a geneticist, we have to study how the expression of the production characters of milk, fat and protein is changed in different zones of Spain (CC, MN, ES).

```{r}
#| echo: true
#| eval: true
#install.packages
#install.packages("tidyverse")
#devtools::install_github("thomasp85/patchwork")
#install.packages("gridExtra")  # Instalar el paquete gridExtra

#if (packageVersion("devtools") < 1.6) {
#  install.packages("devtools")
#}
#devtools::install_github("hadley/lazyeval")
#devtools::install_github("hadley/dplyr")

```
Loading Libraries 
```{r}
#| echo: true
#| eval: true

library(tidyverse)
library(ggplot2)
library(dplyr)
library(plyr)
 
```
Editing the data 
Only we used the data of the 1 lactation of the 3 Zones of Spain 
```{r}
#| echo: true
#| eval: true
rm(list=ls())
### crear un spacio de trabajo
setwd("K:\\Trabajo Maria Jesus\\EstudioGxE\\GxE_files")
# Loading the data
datos_raw<- read.table("Rumlac01_SCS.txt")
dim(datos_raw)
colnames(datos_raw)<-c("R_SC","S_SC","MP","Fat","P","F_p","P_p","Loc","H_Y_S","H_COn_D",
                       "L_Age","Co_DIM","DIM","Animal","hetdimthi",
                       "hetdimthiclass","cov1_thimu012","cov2_thimu012",
                       "cov3_thimu012","cov1_dim","cov2_dim","cov3_dim","CO_Herd",
                       "N_Part_L","D_Control","F_PARTO","dist_t","dist_hr","THI012")
# ####### Redondear la variable THI012 a 2 decimales##############
datos_raw <- datos_raw %>%
  mutate(THI012 = round(THI012, 0))
############ crear la columina zona en España  con solo 3 zonas ############
datos_raw <- datos_raw %>%
  mutate(Zona =case_when(
    Loc=='G'|Loc=='S'|Loc=='J'|Loc=='V'~'CC', # Cornisa cantábrica
    Loc=='L'|Loc=='P'~'MN',  # Mesa norte
    Loc=='C'|Loc=='N'|Loc=='A'~'ES' # España seca
  )) 
########### separar herd-year-season a (Herd y la fecha (que es season_year)) ###########
datos_raw<-datos_raw %>%
  mutate(Herd = str_extract(H_Y_S, "\\d+"),
         fecha = str_replace(H_Y_S, ".*_(\\d{4}_\\d+)$", "\\1"))
################### sperar L_Age a (Lact, la edad (Age)) ######
datos_raw <- datos_raw %>%
  mutate(Lact = case_when(substr(L_Age, 1, 1) == "1" ~ 1,
                          substr(L_Age, 1, 1) == "3" ~ 3,
                          TRUE ~ 2),
         Age = substr(L_Age, 2, 4))
########################### concatenar lact con Zona  ############
datos_raw <- datos_raw %>%
  mutate(Zona_lact = paste(Zona,Lact,  sep = "_"))
head(datos_raw)
#######################################################################################
############################# creat THI clases #############################
datos_raw <- datos_raw %>%
  mutate(THI_c = case_when(
    between(THI012, 40, 41) ~ 1,  between(THI012, 42, 43) ~ 2,  between(THI012, 44, 45) ~ 3,
    between(THI012, 46, 47) ~ 4,  between(THI012, 48, 49) ~ 5,  between(THI012, 50, 51) ~ 6,
    between(THI012, 52, 53) ~ 7,  between(THI012, 54, 55) ~ 8,  between(THI012, 56, 57) ~ 9,
    between(THI012, 58, 59) ~ 10, between(THI012, 60, 61) ~ 11, between(THI012, 62, 63) ~ 12,
    between(THI012, 64, 65) ~ 13, between(THI012, 66, 67) ~ 14, between(THI012, 68, 69) ~ 15,
    between(THI012, 70, 71) ~ 16, between(THI012, 72, 73) ~ 17, between(THI012, 74, 75) ~ 18,
    between(THI012, 76, 77) ~ 19, between(THI012, 78, 79) ~ 20, between(THI012, 80, 81) ~ 21,
    TRUE ~ NA_integer_  ))
########################## concatinar la Zone y THI_c ####################
datos_raw <- datos_raw %>%
  mutate(THI_Zona = paste(Zona,THI_c,  sep = "_"))
############################################################################
################ create MP as a different traits en each zone ############
datos_raw<- datos_raw %>%
  mutate(
    MP_CC = MP,
    MP_MN = MP,
    MP_ES = MP
  )
################### cambios condicionales en las columnas MPdepende la zona #####
# in case CC
datos_raw <- datos_raw %>%
  mutate(
    MP_MN = ifelse(Zona == "CC", 0, MP_MN),
    MP_ES = ifelse(Zona == "CC", 0, MP_ES),
    MP_CC = ifelse(Zona == "MN", 0, MP_CC),# in case MN
    MP_ES = ifelse(Zona == "MN", 0, MP_ES),
    MP_MN = ifelse(Zona == "ES", 0, MP_MN),# in case ES
    MP_CC = ifelse(Zona == "ES", 0, MP_CC)
  ) 
############################################################################
################ create fat as a different traits en each zone ############
datos_raw<- datos_raw %>%
  mutate(
    Fat_CC = Fat,
    Fat_MN = Fat,
    Fat_ES = Fat
  )
################### cambios condicionales en las columnas MPdepende la zona #####
# in case CC
datos_raw <- datos_raw %>%
  mutate(
    Fat_MN = ifelse(Zona == "CC", 0, Fat_MN),
    Fat_ES = ifelse(Zona == "CC", 0, Fat_ES),
    Fat_CC = ifelse(Zona == "MN", 0, Fat_CC),# in case MN
    Fat_ES = ifelse(Zona == "MN", 0, Fat_ES),
    Fat_MN = ifelse(Zona == "ES", 0, Fat_MN),# in case ES
    Fat_CC = ifelse(Zona == "ES", 0, Fat_CC)
  )
############################################################################
################ create protein as a different traits en each zone ############
datos_raw<- datos_raw %>%
  mutate(
    P_CC = P,
    P_MN = P,
    P_ES = P
  )
################### cambios condicionales en las columnas MPdepende la zona #####
# in case CC
datos_raw <- datos_raw %>%
  mutate(
    P_MN = ifelse(Zona == "CC", 0, P_MN),
    P_ES = ifelse(Zona == "CC", 0, P_ES),
    P_CC = ifelse(Zona == "MN", 0, P_CC),# in case MN
    P_ES = ifelse(Zona == "MN", 0, P_ES),
    P_MN = ifelse(Zona == "ES", 0, P_MN),# in case ES
    P_CC = ifelse(Zona == "ES", 0, P_CC)
  )
############################################################################
################ create fat as a different traits en each zone ############
datos_raw<- datos_raw %>%
  mutate(
    F_p_CC = F_p,
    F_p_MN = F_p,
    F_p_ES = F_p
  )
################### cambios condicionales en las columnas MPdepende la zona #####
# in case CC
datos_raw <- datos_raw %>%
  mutate(
    F_p_MN = ifelse(Zona == "CC", 0, F_p_MN),
    F_p_ES = ifelse(Zona == "CC", 0, F_p_ES),
    F_p_CC = ifelse(Zona == "MN", 0, F_p_CC),# in case MN
    F_p_ES = ifelse(Zona == "MN", 0, F_p_ES),
    F_p_MN = ifelse(Zona == "ES", 0, F_p_MN),# in case ES
    F_p_CC = ifelse(Zona == "ES", 0, F_p_CC)
  )
############################################################################
################ create protein as a different traits en each zone ############
datos_raw<- datos_raw %>%
  mutate(
    P_p_CC = P_p,
    P_p_MN = P_p,
    P_p_ES = P_p
  )
################### cambios condicionales en las columnas MPdepende la zona #####
# in case CC
datos_raw <- datos_raw %>%
  mutate(
    P_p_MN = ifelse(Zona == "CC", 0, P_p_MN),
    P_p_ES = ifelse(Zona == "CC", 0, P_p_ES),
    P_p_CC = ifelse(Zona == "MN", 0, P_p_CC),# in case MN
    P_p_ES = ifelse(Zona == "MN", 0, P_p_ES),
    P_p_MN = ifelse(Zona == "ES", 0, P_p_MN),# in case ES
    P_p_CC = ifelse(Zona == "ES", 0, P_p_CC)
  )
######################
datos_raw <- datos_raw %>%
  mutate(
    THI = case_when(
      Zona == "CC" ~ THI012 - 67.86,
      Zona == "MN" ~ THI012 - 69.5,
      Zona == "ES" ~ THI012 - 75.04,
      TRUE ~ THI012
    )
  )
# Loading pedigree
Ped<- read.table("ped.dat", h=TRUE)
colnames(Ped)<-c("Animal","sire","dam")
dim(Ped)
######################### merge the data with pedigree file ###################
datos <- merge(datos_raw,Ped,  by = "Animal", all.x = TRUE)
dim(datos_raw)
dim(datos)
################  Escribir el objeto datos_raw en un archivo de texto para los analysis ###################################
dat <- datos %>%
         dplyr::select(
                MP_CC, MP_MN, MP_ES, Fat_CC, Fat_MN, Fat_ES,
                P_CC, P_MN, P_ES, F_p_CC, F_p_MN, F_p_ES,
                P_p_CC, P_p_MN, P_p_ES, Lact, H_Y_S, L_Age,
                Zona_lact, Age, THI, sire, Animal, THI_Zona, THI_c,
                THI012, cov1_thimu012, cov2_thimu012, cov3_thimu012, cov1_dim,
                cov2_dim, cov3_dim, CO_Herd, R_SC, S_SC, Loc,
                H_COn_D, Co_DIM, DIM, hetdimthi, hetdimthiclass,
                N_Part_L, D_Control, F_PARTO, Zona, Herd, fecha
                ) 

write.table(dat,"data.txt",quote=FALSE,col.names=FALSE,row.names=FALSE)
```